\mythesistitleNOTSYNPOSIS

### 2013-03-21 src/response_to_reviewers.tex

\documentclass{article}
\usepackage{../lib/mythesislib}
\usepackage{../lib/mylatexlib}
\begin{document}

\title{Response to M.S. Thesis Reviewers\\ 
 {\em Thesis Title ``Generalization of the Consecutive-ones Property''}}
\author{Anju Srinivasan}
\date{25 March 2013}

\maketitle

\section{Reviewers}

\begin{enumerate}
\item  {\bf Dr. C. R. Subramaniam}\\ Institute of Mathematical Sciences, Chennai
\item  {\bf Dr. Subhas Chandra Nandy}\\ Indian Statistical Institute, Kolkata
\end{enumerate}

\section{Response to Reviewer 1}

\subsection{Observations}
The following are a few observations mentioned in the review document
with my responses/changes.

\begin{enumerate}
\item Observation: (in Page 1, second last paragraph) {\bf The result on characterization of TPL in
    the thesis is a variant of the 1978 JCT(B) result due to Fournier
  who showed that two hypergraphs $H_1 = (V_1, X)$ and $H_2 = V_2, Y$
such that $|X| = |Y| = m$ are isomorphic iff there is a bijection
$\phi$ on $I = \{1, ... , m\}$ s.t. for every $L \subset I,
|\bigcap_{i \in L}X_i| = |\bigcap_{j=\phi(i),i \in L}Y_j|$} \\
Response: This is an important observation and was not cited in the thesis. Also \cite{f80}
generalizes \cite{br72} and \cite{fg65} by characterizing the
isomorphism of two hypergraphs by means of equicardinality of certain
edge intersections and the exclusion of certain pairs of
subhypergraphs. TPL characterization is a special case of this with
one of the hypergraphs having hyperedges that are paths from a tree
and the characterization only uses edge intersections of at most 3
hyperedges. This has been added in the Conclusions chapter Page 72
under {\em Graph isomorphism and logspace canonization}.

\item Observation: (in Page 2, first paragraph) {\bf A $k$-subdivided tree is
  a tree with a single central vertex with a number of paths each of
  $k$ edges emanating from it.}\\
Response: $k$-subdivided as defined in this thesis (Section
3.2.4, Page 43) is such that each
of the aforementioned paths are of length ``$k+1$'' edges since each such
path has $k$ nodes apart from the central node and the leaf node, i.e.
$k+2$ nodes. No change in the thesis.
\end{enumerate}

\subsection{Comments}
Responses to the comments in review document Page 2 items 0 to 7.

\begin{enumerate}
\item[0] Comment: {\bf In thesis Page 7, Definition 1.3.7: After $x_m
  \le a$, you need to add $x_m \ne a$.}\\ 
Response: True, earlier definition did not accomodate the fact $x_m
\le x_m$. Suggested change made in Page 7 to the definiton of Poset.

\item Comment: {\bf In thesis Page 7, Definition
    1.3.8: $E$ is a set of ``unordered'' pairs. You need to add the
    word ``unordered''}\\
Response: Suggested change made in to definition of Undirected Graph Page 7 Definition 1.3.8 (1).

\item Comment: {\bf In thesis Page 8, A tree is a {\em connected} and
    acyclic graph}\\
Response: The word ``connected'' has been added to the definition of Tree in
Page 7 Definition 1.3.8 (3).

\item Comment: {\bf In thesis Page 8, Definition 1.3.10 you need to
    add that one is looking at intervals of integers. Also define
    isomorphism between graphs for the sake of completeness}.\\
Response: Definition of Interval graph has been changed to say ``a set
of intervals (of integers)'' in Page 8, Definition 1.3.10. Graph
isomorphism definition has been added in Page 7, Definition 1.3.8(4)
as follows: A \textbf{graph isomorphism} between two graphs $G_1$ and
    $G_2$ is a bijection $\phi: V(G_1) \rightarrow V(G_2)$ such that
    $u, v \in V(G_1)$ are adjacent in $G_1$ if and only if $\phi(u),
    \phi(v)$ are adjacent in $G_2$. If such a bijection exists, it is
    said that $G_1$ is \textbf{isomorphic} to $G_2$, denoted by $G_1
    \cong G_2$.

\item Comment: {\bf In thesis Page 8, Definition 1.3.10 definition of
path graph: You need to emphasise that $\cP$ forms a set system}\\
Response: This change has been made - ``a set system of paths $\cP$''
in Page 8 Definition 1.3.10(2).

\item Comment: {\bf In thesis Page 12, third para of Section 1.5: you
    need to write it more precisely and carefully conveying the
    idea. As it is it is not clear what you are saying.}\\
  Response: The problem {\sc Compute feasible tree path labeling} and
  its result have been described formally and precisely in Chapter 3
  (New results). The aforementioned paragraph describes the same
  informally. These informal sentences have been footnoted with the
  corresponding definitions and theorems from Chapter 3. This change
  is in Page 12.

\item Comment: {\bf In thesis Page 28, Definition 7 (of binary
    operator /): it is not defined precisely. What do you mean by
    ``$b$ is a new element added..''?}.\\
Response: This was in thesis Page 26. $A/B$ is created by removing the elements of second operand set $B$ from
the first operand set $A$ and introducing a new element {\em not} in the
universe to $A$ to represent set $B$. This is a set operation 
mentioned in \cite{mm96} (they use ``$B$'' to denote the new element to represent
set $B$ which is also confusing) which is eventually used to decompose
permutation of elements in $A$ to permutation of elements in $A/B$ and
permutation of elements in $B$.
Perhaps it would be clearer if
$b$ is replaced by $x_B$. This change has been made in Page 26
Definition 2.2.4 (7) and other mentions of this operation ``/'' Page
27 item (iii) and $h$ has been replaced with $x_H$ in Corollary
2.2.8. There was also a typo in Page 27 paragraph under Corollary
2.2.8 item (ii) where ``$U \setminus H$'' was written instead of $U / H$.
This has also been corrected.

\item Comment: {\bf In thesis Page 39 Definition 2.3.1: the definition
  of $A_\cF$ is not clear. Where is the dependence on $\cF$?}\\
Response: This was in thesis Page 36. It is true, there is no
dependence in $\cF$. It is only dependent on the universe $U$ of
$\cF$. All occurences of $A_\cF$ has been changed to $A_\U$ in Page 36.
\end{enumerate}

\section{Response to Reviewer 2}

\bibliographystyle{alpha}%
\bibliography{../lib/cop-variants__thesis} %

\end{document}

### 2013-03-21 src/010chap1.tex
Reviewer 1 (CRS) changes made.

  \begin{enumerate}
  \hangindent \defindent
%  \item \textbf{hasse diagram}
  \item An element $x_m \in X$ is a
    \textbf{maximal upper bound} of $X$ if $\nexists a \in X$ such
    that $x_m \posetr a$.  A maximal upper bound on $X$ is
    denoted by $mub(X)$.

@@

\begin{definition}[Graph, Tree, Path]
  \label{def:graphtree}
  \begin{enumerate}
    \hangindent \defindent
  \item An \textbf{(undirected) graph} is $G = (V,E)$ is \stt $V$ is a
    finite set and $E$ is a set of ordered pairs, $E \subseteq V
    \times V$. An element in $V$ is called a \textbf{vertex
      (\emph{pl.} vertices)} and an element from $E$ is called an
    \textbf{edge}. $V$ and $E$ may be written as $V(G)$ and $E(G)$
    respectively. If $u,v \in V$ and $(u,v) \in E$ then $v$ is said to
    be \textbf{adjacent} to $u$ and vice versa. A \textbf{subgraph} of
    $G$ is a graph $G' = (V',E')$ \stt $V' \subseteq V$ and $E'
    \subseteq E$.

@@

  \item A \textbf{tree} is a graph $T$ that has no subgraph that is a
    cycle.
  \end{enumerate}
\end{definition}

@@

\begin{definition}[Interval graph, Path graph]
  \label{def:pathgraph}
  Let $G$ be a graph.
  \begin{enumerate}
  \hangindent \defindent
\item $G$ is an \textbf{interval graph} if there exists a set of
  intervals $\cI$ \stt $G$ is isomorphic to the intersection graph of
  $\cI$. \[G \cong \bI(\cI)\]
  \item $G$ is a \textbf{path graph} if there exists a (undirected
    unrooted) tree $T$ and a set of paths $\cP$ from $T$ \stt $G$ is
    isomorphic to the intersection graph of $\cP$. \[G \cong
    \bI(\cP)\] Path graphs are also known as \uvgraphs (Undirected
    tree where paths intersect in a Vertex) \cite{mw86}.
  \item $G$ is a \textbf{chordal graph} if it has no induced cycle of
    length greater than 3. Moreover, a chordal graph is characterized
    as an intersection graph of subtrees of a tree.
  \end{enumerate}
\end{definition}

@@

The problem of interest in this thesis, namely, tree path labeling
problem, is a natural generalization of the interval assignment
problem or the COT problem. The problem is defined as follows -- given
a set system $\cF$ from a universe $U$ and a target tree $T$, does
there exist a bijection from $U$ to the vertices of $T$ such that each
set in the system maps to a path in $T$.  We refer to this as the
{\CFTPL} problem or simply {\em tree path labeling} problem for an
input set system and target tree pair -- $(\cF,T)$. The special case
of the target tree being a path, is the interval assignment problem.
We focus on generalizing the notion of an ICPIA \cite{nsnrs09} to
characterize feasible path assignments.  We show that for a given set
system $\cF$, a tree $T$, and an assignment of paths from $T$ to the
sets, there is a feasible\footnote{The notion of {\em feasibility} is
  formally defined in Section~\ref{ch:prelims}.}  bijection between
$U$ and $V(T)$ if and only if the intersection cardinalities among any
three sets (not necessarily distinct) is equal to that of the
corresponding paths assigned to them and the input passes a filtering
algorithm (described in this paper) successfully.  This algorithmic
characterization gives a natural data structure that stores all the
feasible bijections between $U$ and $V(T)$. This reduces the search
space for the solution considerably from the universe of all possible
bijections between $U$ and $V(T)$ to only those bijections that
maintain the characterization.  Further, the filtering algorithm is
also an efficient algorithm to test if a tree path
labeling\footnote{The terms {\em tree path labeling} and {\em tree
    path assignment} are, in informal language, synonyms. Formally,
  the former refers to the bijection $\cl: \cF \rightarrow \cP$. The
  latter refers to the set of ordered pairs $\{(S, P) \mid S \in \cF,
  P \in \cP\}$. $\cP$ is a set of paths on $T$.} is feasible.

>>>>

  \begin{enumerate}
  \hangindent \defindent
%  \item \textbf{hasse diagram}
  \item An element $x_m \in X$ is a
    \textbf{maximal upper bound} of $X$ if $\nexists a \in X$ such
    that $x_m \ne a$ and $x_m \posetr a$.  A maximal upper bound on $X$ is
    denoted by $mub(X)$.

@@

\begin{definition}[Graph, Tree, Path, Graph Isomorphism]
  \label{def:graphtree}
  \begin{enumerate}
    \hangindent \defindent
  \item An \textbf{(undirected)} \textbf{graph} is $G = (V,E)$ is \stt $V$ is a
    finite set and $E$ is a set of unordered pairs, $E \subseteq V
    \times V$. An element in $V$ is called a \textbf{vertex
      (\emph{pl.} vertices)} and an element from $E$ is called an
    \textbf{edge}. $V$ and $E$ may be written as $V(G)$ and $E(G)$
    respectively. If $u,v \in V$ and $(u,v) \in E$ then $v$ is said to
    be \textbf{adjacent} to $u$ and vice versa. A \textbf{subgraph} of
    $G$ is a graph $G' = (V',E')$ \stt $V' \subseteq V$ and $E'
    \subseteq E$.

@@

  \item A \textbf{tree} is a connected graph $T$ that has no subgraph that is a
    cycle.

  \item A \textbf{graph isomorphism} between two graphs $G_1$ and
    $G_2$ is a bijection $\phi: V(G_1) \rightarrow V(G_2)$ such that
    $u, v \in V(G_1)$ are adjacent in $G_1$ if and only if $\phi(u),
    \phi(v)$ are adjacent in $G_2$. If such a bijection exists, it is
    said that $G_1$ is \textbf{isomorphic} to $G_2$, denoted by $G_1
    \cong G_2$.
  \end{enumerate}
\end{definition}

@@

\begin{definition}[Interval graph, Path graph]
  \label{def:pathgraph}
  Let $G$ be a graph.
  \begin{enumerate}
  \hangindent \defindent
\item $G$ is an \textbf{interval graph} if there exists a set of
  intervals (of integers) $\cI$ \stt $G$ is isomorphic to the intersection graph of
  $\cI$. \[G \cong \bI(\cI)\]
  \item $G$ is a \textbf{path graph} if there exists a (undirected
    unrooted) tree $T$ and a set system of paths $\cP$ from $T$ \stt $G$ is
    isomorphic to the intersection graph of $\cP$. \[G \cong
    \bI(\cP)\] Path graphs are also known as \uvgraphs (Undirected
    tree where paths intersect in a Vertex) \cite{mw86}.
  \item $G$ is a \textbf{chordal graph} if it has no induced cycle of
    length greater than 3. Moreover, a chordal graph is characterized
    as an intersection graph of subtrees of a tree.
  \end{enumerate}
\end{definition}

@@

The problem of interest in this thesis, namely, tree path labeling
problem, is a natural generalization of the interval assignment
problem or the COT problem. The problem is defined as follows -- given
a set system $\cF$ from a universe $U$ and a target tree $T$, does
there exist a bijection from $U$ to the vertices of $T$ such that each
set in the system maps to a path in $T$.  We refer to this as the
{\CFTPL} problem \footnote{Problem is formally defined in
  Section~\ref{sec:myresearchintro}.} or simply {\em tree path
  labeling} problem for an input set system and target tree pair --
$(\cF,T)$. The special case of the target tree being a path, is the
interval assignment problem.  We focus on generalizing the notion of
an ICPIA \cite{nsnrs09} to characterize feasible path assignments.  We
show that for a given set system $\cF$, a tree $T$, and an assignment
of paths from $T$ to the sets, there is a feasible\footnote{The notion
  of {\em feasibility} is formally defined in
  Section~\ref{ch:prelims}.}  bijection between $U$ and $V(T)$ if and
only if the intersection cardinalities among any three sets (not
necessarily distinct) is equal to that of the corresponding paths
assigned to them and the input passes a filtering algorithm (described
in this thesis) successfully\footnote{Formally defined by
  Theorem~\ref{th:charac}}.  This algorithmic characterization gives a
natural data structure that stores all the feasible bijections between
$U$ and $V(T)$. This reduces the search space for the solution
considerably from the universe of all possible bijections between $U$
and $V(T)$ to only those bijections that maintain the
characterization.  Further, the filtering algorithm is also an
efficient algorithm to test if a tree path labeling\footnote{The terms
  {\em tree path labeling} and {\em tree path assignment} are, in
  informal language, synonyms. Formally, the former refers to the
  bijection $\cl: \cF \rightarrow \cP$. The latter refers to the set
  of ordered pairs $\{(S, P) \mid S \in \cF, P \in \cP\}$. $\cP$ is a
  set of paths on $T$.} is feasible.

### 2013-03-21 src/020chap2.tex
Reviewer 1 (CRS) changes made.

  \item \label{def::slashop} The binary operator $/$ is defined as
    follows. Let $A, B$ be two sets from set system $\cF$.
    \begin{align*}
      A/B &= A \setminus B \cup \{b\}\\
      \cF/B &= \{ S/B \mid S \in \cF, S \nsubseteq B \}
    \end{align*}
    $b$ is a new element added as a representative for $B$.
 \end{enumerate}
  \dstop
\end{definition}

@@

\def \ff {\overline{\cF} \cap \cF^{\bot}}
The idea behind Theorem~\ref{th:mmdecomp} is as follows. A permutation $\alpha$
of $U$ is a composition of two permutations with respect to $H$ - (i) a
permutation $\gamma$ of $H$ and (ii) a permutation $\beta$ of $U
\setminus H$.

For two mutually orthogonal sets $A, B$ such that $A \nsubseteq B$,
$A/B$ is defined as the set obtained by removing all elements of $B$
from $A$ and adding a representative element for $B$ in $A$. Being orthogonal,
$A/B$ results in only the following three possibilities.
\begin{enumerate}[i. ]
\item $A$ and $ B$ are disjoint: $A/B  = A$
\item $A$ is a subset of $B$: $A/B$ is not defined
\item $B$ is a subset of $A$: $A/B = A \setminus B \cup \{b\}$, where $b$
is a new element not in $U$ added to represent $B$.
\end{enumerate}

@@

\begin{corollary}[{\cite[Cor.~8]{mm96}}]
  \label{cor:pqrrecursive}
  Let $\cF$ be a set system with universe $U$ and $H$ is a non-empty
  orthogonal set $H \in \ff$. If there is a \PQRtree $T_1$ that
  encodes all permutations in $\valid{U/H, \cF/H}$ and also a \PQRtree
  $T_2$ that encodes all permutations $\valid{H,\cF \cap \power{H}}$,
  then a \PQRtree $T$ for $\cF$ can be obtained by replacing the leaf
  $h$ in $T_1$ by $T_2$, where $h$ is the representative element for
  $H$ in $T_i$.
\end{corollary}

@@

The work of this thesis described in Chapter~\ref{ch:myresearch} is
largely an extension of the ICPIA characterization. In the new work,
we adopt the notation used in \cite{kklv10} (also described in that
chapter).

@@

\begin{definition}[{Incompatibility graph \cite[Def.~6.1]{mcc04}}]
  \label{def:incompatibility}
  Let $\cF$ be a set system with universe $U$. Let $A_{\cF} = \{(a,b)
  \mid a, b \in U, a \ne b\}$. The {\em incompatibility graph}
  $G_I^\cF$ of $\cF$ is an undirected graph defined as follows.
%  \singlespacing
  \begin{enumerate}
  \item   The vertex set of $G_I^\cF$ is $ A_{\cF}$
  \item  The edge set of $G_I^\cF$ are pairs of the following forms
    \begin{itemize}
    \item $\{(a,b),(b,a)\}$ for all $a, b \in U$
    \item $\{(a,b),(b,c)\}$ if there exists $S \in \cF$ \stt $a, c
      \in S$ and $b \notin S$, for all $a, b, c \in U$
    \end{itemize}
  \end{enumerate}
  \dstop
\end{definition}

Thus an edge between two vertices $(a,b), (b,c)$ in $G_I^{\cF}$ means
that there exists no linear order on $U$ that can place $b$ after $a$
and place $c$ after $b$.  For any \COP order, there must not be any
incompatible pairs. Thus it must consist of an independent set $I$ of
the incompatibility graph. Moreover the independent set must have
exactly half the vertex set of $G_I^\cF$ since the \COP order must
involve all elements in $U$ and each vertex denotes a pair of elements. Secondly, the reverse of a \COP order is
also a \COP order. Thus $A_\cF \setminus I$ must also be an
independent set. In other words, $G_I^\cF$ must be bipartite with
partitions being of equal size if $\cF$ has \COP. An odd cycle is a
forbidden structure in a bipartite graph, hence the same is forbidden
in the incompatibility graph of a set system with \COP. In other
words, if the set system has no \COP, its incompatibility graph must
have an odd cycle.  There is also a requirement on the length of this
odd cycle which is formally stated in Theorem~\ref{th:mccbipartite}.

>>>>

  \item \label{def::slashop} The binary operator $/$ is defined as
    follows. Let $A, B$ be two sets from set system $\cF$.
    \begin{align*}
      A/B &= A \setminus B \cup \{x_B\}\\
      \cF/B &= \{ S/B \mid S \in \cF, S \nsubseteq B \}
    \end{align*}
    $x_B \notin U$ is a new element introduced as shown above as a
    representative for $B$ after removing elements of $B$.
 \end{enumerate}
  \dstop
\end{definition}

@@

\def \ff {\overline{\cF} \cap \cF^{\bot}}
The idea behind Theorem~\ref{th:mmdecomp} is as follows. A permutation $\alpha$
of $U$ is a composition of two permutations with respect to $H$ - (i) a
permutation $\gamma$ of $H$ and (ii) a permutation $\beta$ of $U / H$.

For two mutually orthogonal sets $A, B$ such that $A \nsubseteq B$,
$A/B$ is defined as the set obtained by removing all elements of $B$
from $A$ and adding a representative element for $B$ in $A$. Being orthogonal,
$A/B$ results in only the following three possibilities.
\begin{enumerate}[i. ]
\item $A$ and $ B$ are disjoint: $A/B  = A$
\item $A$ is a subset of $B$: $A/B$ is not defined
\item $B$ is a subset of $A$: $A/B = A \setminus B \cup \{x_B\}$, where $x_B$
is a new element not in $U$ added to represent $B$.
\end{enumerate}

@@

\begin{corollary}[{\cite[Cor.~8]{mm96}}]
  \label{cor:pqrrecursive}
  Let $\cF$ be a set system with universe $U$ and $H$ is a non-empty
  orthogonal set $H \in \ff$. If there is a \PQRtree $T_1$ that
  encodes all permutations in $\valid{U/H, \cF/H}$ and also a \PQRtree
  $T_2$ that encodes all permutations $\valid{H,\cF \cap \power{H}}$,
  then a \PQRtree $T$ for $\cF$ can be obtained by replacing the leaf
  $x_h$ in $T_1$ by $T_2$, where $x_H$ is the representative element for
  $H$ in $T_i$.
\end{corollary}

@@

The work of this thesis described in Chapter~\ref{ch:myresearch} is
largely an extension of the ICPIA characterization. In the new work,
we adopt the notation used in \cite{kklv10} namely, that of hypergraph
isomorphism (also described in that chapter).

@@

\begin{definition}[{Incompatibility graph \cite[Def.~6.1]{mcc04}}]
  \label{def:incompatibility}
  Let $\cF$ be a set system with universe $U$. Let $A_U = \{(a,b)
  \mid a, b \in U, a \ne b\}$. The {\em incompatibility graph}
  $G_I^\cF$ of $\cF$ is an undirected graph defined as follows.
%  \singlespacing
  \begin{enumerate}
  \item   The vertex set of $G_I^\cF$ is $A_U$
  \item  The edge set of $G_I^\cF$ are pairs of the following forms
    \begin{itemize}
    \item $\{(a,b),(b,a)\}$ for all $a, b \in U$
    \item $\{(a,b),(b,c)\}$ if there exists $S \in \cF$ \stt $a, c
      \in S$ and $b \notin S$, for all $a, b, c \in U$
    \end{itemize}
  \end{enumerate}
  \dstop
\end{definition}

Thus an edge between two vertices $(a,b), (b,c)$ in $G_I^{\cF}$ means
that there exists no linear order on $U$ that can place $b$ after $a$
and place $c$ after $b$.  For any \COP order, there must not be any
incompatible pairs. Thus it must consist of an independent set $I$ of
the incompatibility graph. Moreover the independent set must have
exactly half the vertex set of $G_I^\cF$ since the \COP order must
involve all elements in $U$ and each vertex denotes a pair of elements. Secondly, the reverse of a \COP order is
also a \COP order. Thus $A_U \setminus I$ must also be an
independent set. In other words, $G_I^\cF$ must be bipartite with
partitions being of equal size if $\cF$ has \COP. An odd cycle is a
forbidden structure in a bipartite graph, hence the same is forbidden
in the incompatibility graph of a set system with \COP. In other
words, if the set system has no \COP, its incompatibility graph must
have an odd cycle.  There is also a requirement on the length of this
odd cycle which is formally stated in Theorem~\ref{th:mccbipartite}.

### 2013-03-21 src/040chap4.tex
Reviewer 1 (CRS) changes made.

How the work in this thesis connects to graph isomorphism and
canonization is as follows. For a matrix with the \COP, the
intersection graph of the input hypergraph is an interval graph.  A
similar connection exists for \TPL to path graphs which is a subclass
of chordal graphs and a superclass of interval graphs. Graphs which
can be represented as the intersection graph of paths in a tree are
path graphs (as seen in Section~\ref{sec:basicprelim}).  % In TPL, the
% intersection graph of the input hypergraph must be a path graph.  % Our
% results have close connections to recognition of path graphs and to
% path graph isomorphism.  
Thus, a hypergraph $\cF$ can be labeled with paths in a tree (path
labeling as defined in Section~\ref{ch:prelims}), \iff the
intersection graph $\bI(\cF)$ is a path graph. Path graphs are a
subclass of chordal graphs since chordal graphs are characterized as
the intersection graphs of subtrees of a tree \cite[Ch.~4,
Sec.~5]{mcg04}. Chordal graphs are of great significance, are
extensively studied, and have several applications. For every chordal
graph, there exists a tree and a family of subtrees of this tree such
that the intersection graph of this family is isomorphic to the
chordal graph \cite{jrw72phd,gav74b,bun74}.  These trees when in a
certain format, are called {\em clique trees} \cite{ppy92} of the
chordal graph. This is a compact representation of the chordal
graph. There has also been work done on the generalization of clique
trees to clique hypergraphs \cite{km02}.  Path graphs are well studied
in the literature (characterization for \uvgraphs was perhaps first
mentioned by \cite{plr70}).
% Clearly, this is a necessary condition in terms of the intersection
% graph of the input hypergraph $\cF$ in \FTPL. However, one can easily
% obtain a counterexample to show the insufficiency of this
% condition\tnote{Give a figure showing insuff.}.
Indeed, path graph recognition is a necessary condition for \CFTPL but
not sufficient.  We know that path graph recognition can be done in
polynomial time \cite[first algorithm.
$O\left(pn^3\right)$]{gav78},\cite[$O\left(p(m+n)\right)$]{aas93}
\cite{sb94}, \cite[linear algorithm]{db95}, \cite[PR-trees]{cha11}
\footnote{$n$ is number of vertices, $m$ is number of edges, $p$ is
  number of maximal cliques in the input graph.}.  Path graph
isomorphism is also known be \gicomplete \cite{bpt96}.  It would be
interesting to pursue research to answer the following questions in
relation to TPL and graph isomorphism.
\begin{enumerate}
\item What does path graph isomorphism hardness tell us about the
  complexity of \CFTPL?
\item Can TPL help in computing a canon for path graphs, especially a
  logspace canon?
\end{enumerate}

>>>>

How the work in this thesis connects to graph isomorphism and
canonization is as follows. For a matrix with the \COP, the
intersection graph of the input hypergraph is an interval graph.  A
similar connection exists for \TPL to path graphs which is a subclass
of chordal graphs and a superclass of interval graphs. Graphs which
can be represented as the intersection graph of paths in a tree are
path graphs (as seen in Section~\ref{sec:basicprelim}).  % In TPL, the
% intersection graph of the input hypergraph must be a path graph.  % Our
% results have close connections to recognition of path graphs and to
% path graph isomorphism.  
Thus, a hypergraph $\cF$ can be labeled with paths in a tree (path
labeling as defined in Section~\ref{ch:prelims}), \iff the
intersection graph $\bI(\cF)$ is a path graph. Path graphs are a
subclass of chordal graphs since chordal graphs are characterized as
the intersection graphs of subtrees of a tree \cite[Ch.~4,
Sec.~5]{mcg04}. Chordal graphs are of great significance, are
extensively studied, and have several applications. For every chordal
graph, there exists a tree and a family of subtrees of this tree such
that the intersection graph of this family is isomorphic to the
chordal graph \cite{jrw72phd,gav74b,bun74}.  These trees when in a
certain format, are called {\em clique trees} \cite{ppy92} of the
chordal graph. This is a compact representation of the chordal
graph. There has also been work done on the generalization of clique
trees to clique hypergraphs \cite{km02}.  Connections between
hypergraph isomorphism and properties of set systems is also explored
in \cite{br72} by extending Whitney's theorem of graph isomorphism to
hypergraphs.  Moreover, \cite{f80} generalizes \cite{br72} and
\cite{fg65} by characterizing the isomorphism of two hypergraphs by
means of equicardinality of certain edge intersections and the
exclusion of certain pairs of subhypergraphs.  Thus ICPPL is also a
variant of the characterization in \cite{f80} with one of the
hypergraphs having hyperedges that are paths from a tree (path
hypergraph) and ICPPL uses only equicardinality of edge intersections
of at most three hyperedges.

Path graphs are well studied
in the literature (characterization for \uvgraphs was perhaps first
mentioned by \cite{plr70}).
% Clearly, this is a necessary condition in terms of the intersection
% graph of the input hypergraph $\cF$ in \FTPL. However, one can easily
% obtain a counterexample to show the insufficiency of this
% condition\tnote{Give a figure showing insuff.}.
Indeed, path graph recognition is a necessary condition for \CFTPL but
not sufficient.  We know that path graph recognition can be done in
polynomial time \cite[first algorithm.
$O\left(pn^3\right)$]{gav78},\cite[$O\left(p(m+n)\right)$]{aas93}
\cite{sb94}, \cite[linear algorithm]{db95}, \cite[PR-trees]{cha11}
\footnote{$n$ is number of vertices, $m$ is number of edges, $p$ is
  number of maximal cliques in the input graph.}.  Path graph
isomorphism is also known be \gicomplete \cite{bpt96}.  It would be
interesting to pursue research to answer the following questions in
relation to TPL and graph isomorphism.
\begin{enumerate}
\item What does path graph isomorphism hardness tell us about the
  complexity of \CFTPL?
\item Can TPL help in computing a canon for path graphs, especially a
  logspace canon?
\end{enumerate}
